### 2012-10-13 manuscript.tex
Integrate comments from paper walkthrough

% AAMAS 2013 Manuscript

% This file should be compiled with "aamas2013.cls" 
% ----------------------------------------------------------------------
% REMEMBER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
% ----------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STYLE NOTES
%
% TENSE AND VOICE
% * Actual experimental events
% 	We wrote descriptions of actual experimental events using past tense, in 
%	active voice, as the first person plural.
%
% * All other contexts
%	We write all other text using present tense, in active voices, as the
%	the first person plural.  
%
% CONVENTIONS
% flapping-wing, small-scale, lightweight, MAV
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SETUP
\documentclass{aamas2013}

%----------------------------------------------------------------------------%
% PACKAGES
\usepackage{color}
\usepackage[pdftex]{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\providecommand{\norm}[1]{\lVert#1\rVert}

%----------------------------------------------------------------------------%
% METADATA
\def \papertitle{Cooperative Control for Narrow Passage Traversal with an 
Ornithopter Micro Air Vehicle and Lightweight Ground Station}
\def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
particle filter, micro air vehicle, ground station, visual servoing, pose 
estimation, flapping wing, ornithopter, biomimetic}
\def \paperterms{Algorithms, Performance, Design, Experimentation}
\def \paperauthors{Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing}

% Set title
\title{\papertitle}

% PDF bookmarks
\definecolor{linkCol}{gray}{0.25}
\hypersetup{
    pdftitle={\papertitle},
    pdfauthor={\paperauthors},
    pdfsubject={\paperterms},
    pdfkeywords={\paperkeywords},
    pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
    colorlinks, linkcolor=linkCol, citecolor=linkCol, urlcolor=linkCol
}

%----------------------------------------------------------------------------%
% pdfLatex clean-up 
\pdfpagewidth=8.5truein
\pdfpageheight=11truein

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHORS

%% Paper number instead of authors
%\numberofauthors{1}
%\alignauthor Paper  580

\numberofauthors{4}
\author{
\alignauthor Ryan C. Julian\\
       \affaddr{Department of EECS}\\
       \affaddr{Univ. of California, Berkeley}\\
       \affaddr{Berkeley, CA 94720}\\
       \email{ryanjulian@berkeley.edu}
\alignauthor Cameron J. Rose\\
       \affaddr{Department of EECS}\\
       \affaddr{Univ. of California, Berkeley}\\
       \affaddr{Berkeley, CA 94720}\\
       \email{c\_rose@eecs.berkeley.edu}
\alignauthor Humphrey Hu\\
       \affaddr{Robotics Institute}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA 15213}\\
       \email{humhu@cmu.edu}
\and
\alignauthor Ronald S. Fearing\\
       \affaddr{Department of EECS}\\
       \affaddr{Univ. of California, Berkeley}\\
       \affaddr{Berkeley, CA 94720}\\
       \email{ronf@eecs.berkeley.edu}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit 
their on-board sensing and computational resources. Ground vehicles have 
less mobility than MAVs, but relaxed size constraints and typically more 
computing power. These specializations present many opportunities for cooperation. In this work, we 
demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
and a lightweight ground station using computer vision. We develop models for the ornithopter, ground station, and cooperative system dynamics. We determine model parameters of the real systems through experimental system identification. Finally, we verify those models using experiments on narrow passage traversal, and arrive at a cooperative system model which accurately predicts the backwards-reachable region for successfully negotiating ornithopther flight through narrow openings.

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon a previous design by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
payload of 2.8 grams. We augment the ornithopter's built-in 
gyroscope-based control with a lightweight ground station design, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a 
real-time motion tracking algorithm on a live video stream. This allows 
the ornithopter to reliably negotiate flight through narrow openings. We demonstrate the performance of the location estimation and navigation  algorithms.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

% Note that the category section should be completed after reference to 
% the ACM Computing Classification Scheme available at
% http://www.acm.org/about/class/1998/.
% A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, 
% performance measures]
\category{I.2.9}{Robotics}{Autonomous Vehicles}
\category{I.2.9}{Robotics}{Distributed Artificial Intelligence}[Multiagent Systems]
\category{I.2.9}{Robotics}{Vision and Scene Understanding}[Motion]

% General terms should be selected from the following 16 terms: 
% Algorithms, Management, Measurement, Documentation, Performance, 
% Design, Economics, Reliability, Experimentation, Security, 
% Human Factors, Standardization, Languages, Theory, Legal Aspects, 
% Verification.
\terms{\paperterms}

% Keywords are your own choice of terms you would like the paper to be indexed 
% by.
\keywords{\paperkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird ornithopter MAV}
\label{fig:h2bird}
\end{figure}
Power, size, and weight constraints on micro air vehicles (MAV) 
significantly limit sensing their and computational capabilities. They can execute only the most rudimentary
control and state estimation algorithms in real-time. 
The constraints on flapping-wing MAVs are even more restrictive. 
Though they exhibit demonstrate safety and noise profiles superior to those
of rotor craft MAVs, this comes at the expense of reduced payload weight.

The non-linear
flight dynamics exhibited of flapping wing flight compounding constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
Much past research on flapping wing MAVs focused on overcoming the 
difficulties associated with controlling and modeling flapping-wing vehicles in the face of this
non-linearity\cite{floreano:flying}. Common approaches to simplify these 
dynamic models include averaging over the wingbeat period~\cite{Schenato:flightcontrol}
\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict flight behavior~\cite{Shigeoka:ornithopter}. These advances lead to the
development of some high-performance platforms, such as the sub-15 gram DelFly 
described by deCroon et al.~\cite{delfly:design}.

Many control laws that previous researchers developed from
dynamic models of flapping-winged flight, such as periodic forcing and other
control methods described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
require powerful on-board computing. In this work, we develop a simpler, less
computationally intensive method of control for negotiating narrow passages, by exploiting cooperation between specialized robotic agents.

In contrast to flight control, there is considerably less previous work on guidance and of flapping-wing MAVs. 
Baek demonstrated altitude control with an external camera~\cite{baek:altitude}. 
de Croon et al. demonstrated obstacle avoidance with an onboard camera and offboard 
processing~\cite{delfly:avoid}. For sub-20 gram ornithopters, Baek et al. 
\cite{baek:tracking} provides one of the few examples of completely autonomous target seeking for vehicles at this scale.

There also exists a body of work on the guidance and control of fixed- and rotary-wing 
MAV platforms, but many of these methods make extensive use of GPS or motion capture systems
~\cite{kingston:timeattitude}\cite{kanade:3dvision}. Indoors, using
GPS to estimate the pose of an MAV is impossible, and motion capture
requires extensive and stationary modification of the environment. The power
and payload limits on flapping wing MAVs make complex on-board sensing difficult, so multi-agent
cooperative control schemes are apt solutions.

\cite{Jung1998Range}\cite{Hyams1999Cooperative}\cite{Mehta2006Adaptive}
\cite{Rudol2008Micro}\cite{Luo2011Air}

%TODO Add in final paragraph to lead-in to our work

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robotic and Vision Platforms}
%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
To explore these cooperative control concepts, we designed a new flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{h2bird}). Built around the Silverlit i-Bird 
RC flyer power train and clap-fling 
wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Fig.~\ref{h2Bird_axes}. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller consists of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
% ryan
Cooperative behavior is a promising strategy for autonomous navigation of 
unexplored and unstructured environments. However, to operate in these environments,
a cooperative system must be fully mobile, not tied to a static ground station or 
motion capture laboratory. To demonstrate the feasibility of this approach, 
we designed our ground station to conform to the power and weight 
constraints of highly mobile millirobotic ground vehicles, such as 
OctoRoACH, which can traverse rough terrains with speed and agility~\cite{Pullin2012Dynamic}.

Rather than power-intensive PC processors, we used the ARM-based 
BeagleBoard single-board computer as reference computation platform. The
BeagleBoard consumes approximately one watt while running our computer vision 
and control algorithms. It uses the same processor as popular sub-10 gram 
processor modules---such as the Gumstix Overo and LogicPD Torpedo---which are 
light enough for deployment on millirobotic ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird with a
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

Our ground station software is based on the Ubuntu Linux\footnote{\raggedright Ubuntu: 
Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
modified with a custom version of the Linux kernel. We use the Python\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
{http://www.python.org/}} programming language for control, robot communication, 
and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
\href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
computer vision libraries for image capture and processing.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Concept and Algorithms}
%TODO ryan

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/process_flow.pdf}
\caption{Overview of the cooperative control system.}
\label{fig:process_flow}
\end{figure}
%TODO h2bird label in diagram should have superscript

%----------------------------------------------------------------------------%
\subsection{Ornithoper Attitude Control}

\begin{figure}[!tb]
\centering
\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
\caption{H$^2$Bird ornithopter with attitude control axes and labeled control surfaces.}
\label{fig:h2Bird_axes}
\end{figure}
%TODO label h2bird drawing

We implement attitude estimation and control of the H$^2$Bird on-board to achieve 
high sample-rate attitude control. We estimate the vehicle pose by
integrating the IMU, and use a proportional-integral-derivative (PID) 
controller to feed these estimates back to the attitude control flight surfaces (Figs.\ref{process_flow} and~\ref{block_diagram}).

The high pitch angles flapping-wing MAVs regularly operate at distinguish them 
from other MAVs. Baek demonstrated control on a similar vehicle using PID and an Euler angle
parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of vehicle orientation 
suffer from singularities around high pitch angles. So we instead 
represent the vehicle orientation with quaternions. An additional benefit of 
the quaternion representation is its ability to represent relative body and 
world coordinate angular displacements compactly with quaternion multiplication.

Our implementation of quaternion-based control is similar to Knoebe's~\cite{knoebe:quatcontrol}:
Given a reference pose represented in quaternion form $q_r$, we define the error as the rotation $q_e$ 
required to reach the reference pose from the estimated pose $q$ (Eq.~\ref{eq:quat_error}). 
\begin{align}
\label{eq:quat_error}
q_e &= q^{\prime}\cdot q_r \\
		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
\end{align}
We then convert this error quaternion to an angle axis representation and use it 
as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
We perform these calculations on-board, with trigonometric lookup tables for real-time performance.
\begin{align}
\label{eq:quat_linearize}
Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
\end{align}
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}
The ground station performs simple two-dimensional pose estimation on 
H$^2$Bird using a particle filter-based motion tracking algorithm.
The produced pose estimate is the feedback input to the cooperative visual servoing feedback 
loop (Sec. \ref{sec:visual_servoing_concept}). We limit our pose estimation 
approach to monocular two-dimensional object tracking in pixel space. This 
allows us to perform real-time video tracking using the modest computational 
resources available on the ground station, and similarly, millirobotic ground 
vehicles.

Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a particle filter ~\cite{thrun2005probabilistic}. 
We initialize the particles uniformly across the frame before tracking begins. 
To improve numerical stability, we normalize the weights of all particles on 
each iteration, and only resample the particle population when the mean particle 
weight falls below a pre-determined threshold.

%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at time $t$, $\Sigma$ 
is a chosen covariance matrix, and $b$ is vector of the bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a uniform variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across of the pixel space. This helps prevent overconfidence and collapse, 
and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
approach to motion tracking via na\"{i}ve background subtraction, e.g. frame differencing[CITATION].
On each iteration, the system provides the emission model with $f_0$, the frame representing 
the background and $f_t$, the most recent frame captured. The model assigns 
a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
$x^{[i]}_t$.
\begin{equation}
\label{eq:emission_model}
w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
\end{equation} 
Hence, particles are more likely to be resampled in regions of 
the image which are dissimilar in color from the background.

During experiments presented in this work, we assumed that neither the bird, 
nor any other moving object, were visible in the frame on system 
start. We saved the first frame captured by the system, and provided it as 
the background frame $f_0$ for every iteration of the particle filter. This choice of $f_0$ 
resulted in a sampling approach to na\"{i}ve background subtraction.

Na\"{i}ve background subtraction is known to be brittle to changes in the scene over 
time[CITATION]. However, we find it very efficient and sufficiently reliable for the 
short time spans necessary for H$^2$Bird to successfully negotiate narrow 
openings, which are typically less than five seconds. Substituting $f_0$ 
with a more sophisticated background extraction would retrain the underlying motion tracking features of the filter, while providing it with a more robust way of differentiating subject motion from the background.

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by commanding changes to its
reference attitude, as governed by its on-board attitude controllers (Figs.\ref{process_flow} and~\ref{block_diagram}). 
With the estimate of 2D vehicle pose from computer vision, we calculate an altitude
error and horizontal error. The ground station then commands a reference attitude to 
the H$^2$Bird pitch and yaw. according to a PID control law. We apply a fixed-lag window the PID controller integrator term 
to prevent wind-up while we launch the ornithopter. We tuned the proportional and derivative terms of this outer controller 
to dampen the overshoot and drift cause by the cooperative system latency.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Model}
\label{sec:system_model}

%----------------------------------------------------------------------------%
\subsection{Ornithopter Attitude Control}
\label{sec:model_attitude}

We divide our model of the vehicle pose into two independent models for simplicity: one for
yaw motion, and one for translational motion.

We approximate the yaw dynamics of the ornithopter with a one-dimensional model to limit complexity and because it matches our physical implementation. Likewise, we choose the reference heading angle as input to the model and model the dynamics of vehicle and on-board PID controller as one system, which outputs the physical heading angle. We show the resulting model form in Equation~\ref{eq:plain_model}.

\begin{equation}
\label{eq:plain_model}
H(s) = \frac{K}{1 + Rs}
\end{equation}

To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq.~\ref{eq:total_model_plain})~\cite{lavalle:planning}. 
\begin{equation}
\label{eq:total_model_plain}
\begin{aligned}
& \dot{x} = sin(\theta)\\
& \dot{y} = cos(\theta)\\
\end{aligned}
\end{equation}
Our form uses the angular position, $\theta$ as an input rather than a state. We model the overall motion of the system as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (Eq. \ref{eq:plain_model}) with the reference heading angle as the input.

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}

Our model of the ground station comprises models of both the camera and the
pose estimation computation. The camera model contains both a
latency, which is in series with the path between the ``camera'' and ``ornithopter dynamics''
 blocks in Fig.~\ref{fig:block_diagram}, and a model of the camera's viewing angle constraints.

 We model the 
viewing angle of the camera as a pixel map from one edge of a
63$^{\circ}$ viewing triangle to the other. The camera has an image width of
640 pixels. We calculate the desired heading angle in degrees for a given distance from the
camera according to Equation~\ref{eq:desired_heading}, where $d$
is the distance between the ornithopter and the camera, and $x$ is the horizontal
position of the robot relative to the center of the window in meters. This
explicitly fact that the pose estimation algorithm does not estimate the depth of the
robot.
\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
\end{equation}

We model the pose estimation computation by the by the latency of running 
the particle filter, calculating a new reference attitude, and sending to the ornithopter via radio.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}

%----------------------------------------------------------------------------%
\subsection{System Identification}

%............................................................................%
\subsubsection{Ornithoper Attitude Control}

We determined the turning speed and turning radius of the H$^2$Bird
experimentally, by measuring the response of the system to a step input of a
clockwise 90$^{\circ}$ turn. We measured the response of the system by recording
the attitude estimates calculated by the ornithopter using its on-board
gyroscope. We then used
MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} to fit a
simple low-order model to the step response, as discussed in Sec.~\ref{sec:model_attitude}.

%............................................................................%
\subsubsection{Ground Station Pose Estimation}

To measure the capture latency of the camera, we wrote a simple test 
application which changes a large colored rectangle on the ground station's screen, then uses the camera to detect that change. By noting the value of the system timer just before we change the screen, and just after the camera detects that change, the application arrives at an estimate of the video capture pipeline latency. We averaged these result over several hundred rapidly-executed trials to determine the capture latency we used to model the system..

To compute the particle filter and controller latency, we again used system timers to determine the
average time between the start and end of computation.

%----------------------------------------------------------------------------%
\subsection{Model Verification}

We tested our cooperative visual servoing system with a simple target seeking
experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
a wooden frame we used to simulate a window. As the H$^2$Bird cannot detect the
target on its own, we used the cooperative control with the ground station, which to provided remote guidance to H$^2$Bird to allow it to navigate to the window.
We used a motion capture system to record 
ground truth data. We present a sketch of this experimental setup in
Fig.~\ref{fig:experiment_cartoon}.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
\caption{Frame sequence from video feed used for tracking. Particles 
are shown in \textcolor{red}{red}, the current state estimate as a 
\textcolor{yellow}{yellow} circle, the window bounding box in
\textcolor{blue}{blue}, and the target location as a \textcolor{green}{green} 
circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
Particles converge on its location; 4.~Release H$^2$Bird, which 
begins cooperative autonomous flight toward the target; 5.~Visual servoing 
controller overshoots the window on the left, and commands a right-turn 
maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
\label{fig:pf_screencap}
\end{figure}

We performed each trial as follows (see Fig.~\ref{fig:pf_screencap}):
\begin{enumerate}
\item The window is identified by the user. 
\item We release H$^2$Bird by hand, in view of the camera at the desired starting grid point.
\item The H$^2$Bird attempts to fly through the window with, with remote guidance from
the ground station.
\end{enumerate}

In total, we ran 80 trials over a variety of starting positions. Our goal for the experiments was to verify the model described in
Secttion~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid
along the edges of the camera view plane, and in 0.05 meter increments from the
window, in an effort to determine the success rate on the edges of the testing space. We
conducted 20 additional trials directly in front of the camera to determine
the success rate near the center of viewing plane. After every 5 trials,
we replaced the ornithopter's battery with a fully charged one.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
\caption{Conceptual sketch of experimental setup.}
\label{fig:experiment_cartoon}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Performance}
\label{sec:performance}
To bound the performance of the cooperative control system, we determine 
the feasible set of initial conditions that result in successful narrow 
passage traversal, using both analytical and empirical methods. We then conducts 
experiments to validate this model.

The following sections detail how we calculated the feasible and infeasible
regions using system identification and present the results of our
experimental trials.

%----------------------------------------------------------------------------%
\subsection{System Identification}

%............................................................................%
\subsubsection{Ornithoper Attitude Control}
\label{sec:flight_control}
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
\caption{H$^2$Bird step response}
\label{fig:step_response}
\end{figure}
%TODO reduce number of sig figs

The results of the step response experiment are described in
Fig.~\ref{fig:step_response}. The first-order model that we fit to the
response is in Equation~\ref{eq:transfer_func}.

\begin{equation}
\label{eq:transfer_func}
H(s) = \frac{0.95}{1+0.62s}
\end{equation}

The step response rise time is 1.4 seconds. The H$^2$Bird flies
at an average forward velocity of 1.2 m/s during our experiments, so we
estimate the minimum turning radius of the robot is 1.07 maters, as described in
Equation~\ref{eq:min_radius}.
\begin{equation}
\label{eq:min_radius}
\begin{aligned}
\text{Minimum radius} & = \frac{360^{\circ}*\text{Rise time}*\text{Flight
speed}}{\text{Final angle}*2*\pi}\\
& = \frac{360^{\circ}*1.4s*1.2\frac{m}{s}}{90^{\circ}*2*\pi} = 1.07 m
\end{aligned}
\end{equation}
In the experiments presented in this work, we choose a flight speed
lower than the maximum speed of the ornithopter, to facilitate more robust
tracking.

%............................................................................%
\subsubsection{Ground Station Pose Estimation}

The latency that we measured for the camera was 25ms, and the measured
partiicle filter computational latency was 12.5ms.

%............................................................................%
\subsubsection{Cooperative Visual Servoing}
\label{sec:visual_servoing}

\begin{figure*}[tb]
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\linewidth]{figures/flight_paths_feasible.pdf}
\caption{Plot of camera field of view and experimental trials within the feasible region.}
\label{fig:flight_paths_feasible}
\end{minipage}
\hfill
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/flight_paths.pdf}
\caption{Plot of camera field-of-view and perimeter experimental trials overlayed.}
\label{fig:flight_paths}
\end{minipage}
\end{figure*}

We present the results of our experimental trials in
Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
initial conditions in the middle of the testing space and in front of the
camera, we achieved a success rate of 80\%. This case is depicted in
Fig.~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminishes
to 50\% because of several factors. On the edges of the camera frame, the yaw
control inputs are large in magnitude, although this is not necessary in all
cases. Since the control algorithm we are using has no notion of depth, it
applies the same control input whether the ornithopter is close to the camera,
where small changes in heading are adequate, or far away from the camera,
where large changes in heading are necessary. Due to this phenomenon, the
controller often over or under-compensates in certain regions, depending upon
the PID tuning and the distance from the camera. Tracking noise can also cause
passage negotiation to fail. Variations in the H$^2$Bird hand launch can cause
the system to respond differently for similar initial conditions. In addition,
the H$^2$Bird battery was changed approximately every 5 tests, adding more
variation to the H$^2$Bird performance by affecting the motor power output.

%----------------------------------------------------------------------------%
\subsection{Model Verification}

\begin{figure*}[tb]
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
\caption{Plot of backwards reachable set for successful window traversal.}
\label{fig:feasible_set}
\end{minipage}
\hfill
\begin{minipage}[b]{0.45\linewidth}
\centering
\centering
\includegraphics[width=\textwidth]{figures/heat_map.png}
\caption{Probability of success for all starting locations. This graphic is 
a result of a Monte Carlo simulation of the reachable set model, in which the 
initial heading of the H$^2$Bird is chosen uniformly at random from $(-90,90)$.}
\label{fig:heat_map}
\end{minipage}
\end{figure*}

We simulated the motion of the ornithopter, beginning from certain positional
and angular initial conditions using several components representing the
experimental layout of our system described in Fig.~\ref{fig:block_diagram}.
For the ``ornithopter dynamics'' block, we used the translational model
described by~\ref{eq:total_model_plain}, with the angular position model
described by~\ref{eq:transfer_func}. We tuned the PID controller for the
simulation so that paths from known positions looked similar to the behavior
we observed in the experimental data. We also account for the camera capture and
pose estimation delays, and the PID controller sampling rate in the
simulation.

We determine the backwards reachable set of initial states for successful
narrow passage traversal both geometrically and in simulation
(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform
grid of 0.05 meter increments within the camera viewing region, and an initial
angular position perpendicular to the narrow passage plane. The geometrically
infeasible region is computed using the minimum turning radius and the
geometry of the camera's field-of-view and is indicated in blue. Both the
regions in red and blue indicate the infeasible region determined in
simulation. The feasible region is highlighted in green. A success consisted
of a path that intersects the window at some point and does not leave the
camera viewing triangle. Fig.~\ref{fig:feasible_set} is validated by our
experimental data, as there are no successes within the geometrically
infeasible region. There are a few successes near the boundary of the
infeasible region determined in simulation, but this phenomenon can be
explained by the results in Fig.~\ref{fig:heat_map}.

Fig.~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
simulating possible paths of the H$^2$Bird during our experiments. For each
point within a 0.2 meter grid within the camera viewing triangle, we simulated
10 trials using an initial heading randomly sampled from a uniform
distribution between -90$^{\circ}$ and 90$^{\circ}$. As shown in the figure,
there is a region in the middle with a 100\% probability of success, which is
a subset of the feasible region presented in Fig.~\ref{fig:feasible_set}.
Along the edges of the camera viewing triangle, however, the probability of
success decreases.

The results of our experimental trials verify the results of the Monte Carlo
simulation of our model. The trials that we conducted in
Fig.~\ref{fig:flight_paths_feasible} correspond to the circled region in
Fig.~\ref{fig:heat_map}. The 80\% success rate that we measured in the
experiments corroborates our calculated probability of success of
approximately 80\%. Additionally, the diminished success rate of 50\% near the
edges of the viewing triangle that we determined through experimentation
supports the reduced probability of success calculated by the simulation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}
%TODO 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
The authors would like to thank Fernando Garcia Bermudez for his 
assistance with the Vicon motion capture system, Andrew Pullin for his 
help with robot photography, and the members of the Biomimetic 
Millisystems Laboratory and the EECS community at the University of 
California, Berkeley for their advice and support.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BIBLIOGRAPHY

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{manuscript}

% ACM needs 'a single self-contained file'!
%\subsection{References}
% Generated by bibtex from your ~.bib file.  Run latex,
% then bibtex, then latex twice (to resolve references)
% to create the ~.bbl file.  Insert that ~.bbl file into
% the .tex source file and comment out
% the command \texttt{{\char'134}thebibliography}.
\end{document}